### 2022-01-05 Manuscript/Manuscript_Eli.Rmd
add results

---
title: "Identification of quality-reflecting text features using Machine Learning Techniques"
author: "Eli-Boaz Clapper"
date: '`r format(Sys.time(), "%d %B, %Y")`'
abstract: |
  Rating written summaries and using the ratings as a fair proxy for text quality is a time-costly procedure that is prone to rater variability. Computerized techniques that aid the automatic scoring of text can aid the procedure. Previous research has tried to identify quality-reflecting text features to be used in automatization techniques, but the statistical models have been subject to overfitting and multicollinearity, leading to imprecise results. This thesis views the issue in a machine learning framework to help aid the rating automatization and to identify quality-reflecting text features. It is shown that.... and .... this implies that ... and ...

output: 
  html_document:
    css: "style.css"
    fig_caption: yes
bibliography: references.bib
link-citations: yes
linkcolor: blue
---

<br><br><br><br><br><br>

@@

<div id = "fig1" class = "space">
```{r GRT, echo=FALSE}
#![Image 1: *Graphic Representation of Thesis*](../Images/Flowchart_Thesis.jpg)
knitr::include_graphics("../Images/Flowchart_Thesis.jpg")
```
  <p class = "caption"><em>Figure 1.</em> Graphical Representation of the Thesis</p>
</div>

@@

<p class = "firstp"> In total, five ML techniques are used to answer the research questions. These five algorithms produce models for which variable importance can be extracted. These techniques include Ridge-, Lasso-, Elastic net-, Linear Principle Component Regression [LPCR], and Multivariate Linear Regression [MLR] of which the first three are available in package `glmnet` with function `glmnet()` and the latter two are available in the `stats` package with functions `lm()` and `prcomp()` provided by R Core Team [@glmnetJerome; @RCoreTeam]. These techniques were chosen because of their ability for regularization or data reduction, which reduces multicollinearity. These techniques are intended to identify the importance magnitude of text features while dealing with multicollinearity and overfitting. MLR is used as a benchmark to compare performance of the other algorithms to. </p>

@@

<div id = "tab3" class = "space">
<p class = "caption">Table 3. <em>Varying and default-deviating hyperparameters</em></p>
<div class = "caption">
| Algorithm    | Parameter    | Interpretation                    | Value                                                 |
|--------------|--------------|-----------------------------------|-------------------------------------------------------|
| Lasso/Ridge  | $\lambda$    | Amount of regularization          | default in `glmnet()`                                 |
| Elastic Net  | $\lambda$    | Amount of regularization          | default in `glmnet()`                                 |
|              | $\alpha$     | Amount of L1-L2 regularization    | default in `glmnet()`                                 |
| LPCR         | $n_{pc}$     | Number of extracted PC's          | 1-4                                                  |

@@

All algorithms are evaluated on three predictive performance metrics that determine their usability in automatically scoring summaries. Those metrics are the proportion of explained variance in the human ratings by the model: $R^2$, the Root of the Mean Squared Error: $RMSE$, and the Mean Absolute Deviation: $MAD$. The algorithms are additionally investigated on which features/principle components are selected as relevant. Relevance is in turn defined for when a regression coefficient does not contain 0 in its 95\% confidence interval. Additionally, for PCR, a text feature is regarded as important if its factor loading $\geq$ .20 on at least one of the selected principle components. Both $MAD$ and $RMSE$ are used as $MAD$ is more robust against outliers. $R^2$ is used to determine how much variance in the ratings can be explained by the used text features. $RMSE$ will be the leading metric to determine predictive performance.

@@

input Results
<br><br>

### Discussion

input discussion

>>>>

---
title: "Identification of quality-reflecting text features using Machine Learning Techniques"
author: "Eli-Boaz Clapper"
date: '`r format(Sys.time(), "%d %B, %Y")`'
abstract: |
  Rating written summaries and using the ratings as a fair proxy for text quality is a time-costly procedure that is prone to rater variability. Computerized techniques that aid the automatic scoring of text can aid the procedure. Previous research has tried to identify quality-reflecting text features to be used in automatization techniques, but the statistical models have been subject to overfitting and multicollinearity, leading to imprecise results. This thesis views the issue in a machine learning framework to help aid the rating automatization and to identify quality-reflecting text features. It is found that the linear model as well as the Elastic Net algorithms perform best, while Lasso performs worst. This is likely because only six predictors were used in this mini-thesis. It is also found that the number of words a summary contains seems to be the best predictor for the rating. Future research should use all features and all algorithms for more conclusive results.

output: 
  html_document:
    css: "style.css"
    fig_caption: yes
bibliography: references.bib
link-citations: yes
linkcolor: blue
---
<p style= "font-size:120%"> keywords: <em> machine learning, summary writing, automatization</em></p>
<br><br><br><br><br><br>

@@

<div id = "fig1" class = "space">
```{r GRT, echo=FALSE}
#![Image 1: *Graphic Representation of Thesis*](../Images/Flowchart_Thesis.jpg)
knitr::include_graphics("../Images/Flowchart_Thesis.jpg")
```
  <p class = "caption"><em>Figure 1.</em> Graphical Representation of the full Thesis</p>
</div>

@@

<p class = "firstp"> In total, five ML techniques are used to answer the research questions. These five algorithms produce models for which variable importance can be extracted. These techniques include Ridge-, Lasso-, Elastic net-, Linear Principle Component Regression [LPCR], and Multivariate Linear Regression [MLR] of which the first three are available in package `glmnet` with function `glmnet()` and the latter two are available in the `stats` package with functions `lm()` and `prcomp()` provided by R Core Team [@glmnetJerome; @RCoreTeam]. `glmnet()` minimizes the following Residual Sum of Squares [RSS]: </p>

<div class = "space">
\begin{align}

RSS_{ElasticNet} &= \frac{1}{N} \sum_{i=1}^{N} (y_i - \beta_0+\boldsymbol{\beta}^T x_i) + \lambda\left[(1-\alpha)\|\boldsymbol{\beta}\|_2^2/2 + \alpha \|\boldsymbol{\beta}\|_1\right] \\

RSS_{Ridge} &= \frac{1}{N} \sum_{i=1}^{N} (y_i - \beta_0+\boldsymbol{\beta}^T x_i) + \lambda(\|\boldsymbol{\beta}\|_2^2/2) \\

RSS_{Lasso} &= \frac{1}{N} \sum_{i=1}^{N} (y_i - \beta_0+\boldsymbol{\beta}^T x_i) + \lambda\| \boldsymbol{\beta}\|_1

\end{align}
</div>

<p class = "firstp">with respect to $\beta_0$ and $\beta^T$  where $\alpha$ is the tuning parameter for elastic net that determines the amount of Lasso or Ridge regression. Setting $\alpha = 0$ or $\alpha = 1$ defines $RSS$ for Ridge and Lasso Respectively. </p>

These techniques were chosen because of their ability for regularization or data reduction, which reduces multicollinearity. These techniques are intended to identify the importance magnitude of text features while dealing with multicollinearity and overfitting. MLR is used as a benchmark to compare performance of the other algorithms to.

@@

<div id = "tab3" class = "space">
<p class = "caption">Table 3. <em>Varying and default-deviating hyperparameters</em></p>
<div class = "caption">
| Algorithm    | Parameter    | Interpretation                    | Value                                                 |
|--------------|--------------|-----------------------------------|-------------------------------------------------------|
| Lasso/Ridge  | $\lambda$    | Amount of regularization          | default in `glmnet()`                                 |
| Elastic Net  | $\lambda$    | Amount of regularization          | default in `glmnet()`                                 |
|              | $\alpha$     | Amount of L1-L2 regularization    | 0.1-0.9                                 |
| LPCR         | $n_{pc}$     | Number of extracted PC's          | 1-4                                                  |

@@

All algorithms are evaluated on three predictive performance metrics that determine their usability in automatically scoring summaries. Those metrics are the proportion of explained variance in the human ratings by the model: $R^2$, the Root of the Mean Squared Error: $RMSE$, and the Mean Absolute Deviation: $MAD$. The algorithms are additionally investigated on which features/principle components are selected as relevant. Relevance is in turn defined for Lasso when the algorithm selects a feature, for MLR when a regression coefficient does not contain 0 in its 95\% confidence interval, for Ridge and Elastic Net when the absolute regression coefficient > .05. Additionally, for PCR, a text feature is regarded as important if its factor loading $\geq$ .20 on at least one of the selected principle components. Both $MAD$ and $RMSE$ are used as $MAD$ is more robust against outliers. $R^2$ is used to determine how much variance in the ratings can be explained by the used text features. $RMSE$ will be the leading metric to determine predictive performance.

@@

<h4> <strong>Predictive performance</strong> </h4>
<p class = "firstp"> The ability of the algorithms to correctly predict new data points was determined by looking at the $R^2$, $MAD$ and $RMSE$ values they produced. The results indicate that the MLR and Elastic Net algorithms performed best $R^2 = 0.57$, $MAD = 1.71$, $RMSE = 2.27$, closely followed by ridge regression, $R^2 = 0.56$, $MAD = 1.74$, $RMSE = 2.28$ and LPCR, $R^2 = 0.50$, $MAD = 1.89$, $RMSE = 2.42$. Lasso performed worst: $R^2 = 0.34$, $MAD = 2.29$, $RMSE = 2.8$. <a href = "#tab4"> Table 4 </a> shows the results
</p>

<div class = "space" id = "tab4">
<p class = "caption">Table 4. <em>Predictive performance of algorithms</em></p>
```{r resmetrics, echo=F}
results <- readRDS("../Results/Results.RData")

datatable(round(results$performance,2), options = list(dom = 't'))
```
</div>

<h4> <strong>Feature selection</strong> </h4>
<p class = "firstp">It was also investigated how often certain features were selected as relevant by the algorithms. <a href = "#fig2"> Figure 2 </a> shows the results.</p>

<div class = "space" id = "fig2">
<p class = "caption"><em>Figure 2.</em> Feature importance</p>
```{r featureimportance, echo=F}
featimp <- sort(results$feature_imp, decreasing = T)
plot(featimp, type = "b", main = "Number of times feature is selected", xlab = "Features", ylab = "Counts")
```
</div>

### 2022-01-05 Script/2 Analyses.Rmd
add results

load in data
```{r}
TextFeatures <- read.csv2(file = "../Data/6 TextFeatures.csv", sep = ",")[,-1] #load in data minus csv-appended column
TextFeatures <- sapply(TextFeatures, as.numeric) #convert to numeric
```

split data into train and test
```{r}
set.seed(6164900) #seed for reproducability
splits <- sample(1:nrow(TextFeatures), size = floor(nrow(TextFeatures)*0.6))

@@

@@

evaluation metrics
```{r}
source("Functions.R")
```

Multilinear Regression. Does not need CV in this case, no parameters to be tuned
```{r}
fit_lm <- lm(Rating ~ ., data = as.data.frame(train))
pred_lm <- predict(fit_lm, newdata = as.data.frame(test))
metrics_lm <- metrics(test[,1], pred_lm) #not bad actually

@@

preprocess for other algorithms
```{r}
K <- 10 #number of CVs
train <- cbind(train, ind = sample(rep(1:K, length.out = nrow(train)))) #indicator variable
traindf <- as.data.frame(train)
testdf <- as.data.frame(test)

@@

principle component analysis algorithm
```{r}
npc <- 1:4

metrics_lpcr <- list()
for(j in npc){
  metric_cv <- matrix(NA, K, 3)
  colnames(metric_cv) <- c("RMSE", "MAD", "R2")
  for(i in 1:K){
    traincv <- traindf[traindf$ind != i,]
    validation <- traindf[traindf$ind == i,]

   metric_cv[i,] <- LPCR(traincv, validation, npc = j)
  }
  metrics_lpcr[[j-(npc[1]-1)]] <- metric_cv
}

names(metrics_lpcr) <- sprintf("NPC%d", npc)

@@

LPCR obtain best condition
```{r}
(eval_lpcr <- lapply(metrics_lpcr, function(x){ #obtain mean and variances of metrics over folds
  apply(x, 2, function(y){
    c(m = mean(y), variance = var(y))
  })
}))

RMSE_lpcr <- sapply(eval_lpcr, function(x){ #obtain RMSE values vector
  x[1,1]
})

RMSE_lpcr #not much difference between 3 and 4
names(which(RMSE_lpcr == min(RMSE_lpcr))) #using 4 PC's gives the lowest RMSE, however it is not much better than when 3 PC's are used, so 3 is preferred. 
```

LPCR obtain metrics. *do not forget to obtain important variables*
```{r}
LPCR(traindf, testdf, npc = 3)
```

>>>>

used libraries
```{r}
library(glmnet)
```

load in data
```{r}
#load in data minus csv-appended column
TextFeatures <- read.csv2(file = "../Data/6 TextFeatures.csv", sep = ",")[,-1]

#convert to numeric
TextFeatures <- sapply(TextFeatures, as.numeric) 
```

split data into train and test
```{r}
#seed for reproducability
set.seed(6164900)

#proportion of complete set dedicated to training set and split into test and training data
prop.train <- 0.6 
splits <- sample(1:nrow(TextFeatures), size = floor(nrow(TextFeatures)*prop.train))

@@

K <- 10 #number of folds

@@

evaluation metrics and other functions used
```{r}
source("Functions.R")
```

###Multilinear Regression. 
Does not need CV in this case, no parameters to be tuned.
```{r}
fit_lm <- lm(Rating ~ ., data = as.data.frame(train)) #train model
pred_lm <- predict(fit_lm, newdata = as.data.frame(test)) #test model
metrics_lm <- metrics(test[,1], pred_lm) #obtain metrics

@@

###lasso
```{r}
set.seed(6164900)
Lassocv <- cv.glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 1, family = 'gaussian', standardize = F,
                    nfolds = K) #setting alpha to 1, runs Lasso

#evaluate cross validation
plot(Lassocv)

#train model
fit.Lasso <- glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 1, family = 'gaussian', standardize = F,
                    lambda = 1.5)

#obtain important predictors
lasmat <- as.matrix(coef(fit.Lasso))
imp_Lasso <- names(lasmat[lasmat != 0,])[-1]

#evaluate model
pred_lasso <- predict(fit.Lasso, newx = test[,2:ncol(test)])
metrics_lasso <- metrics(test[,1], pred_lasso)

@@

###Ridge
```{r}
K <- 10 #number of CVs

set.seed(6164900)
Ridgecv <- cv.glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 0, family = 'gaussian', standardize = F,
                    nfolds = K) #setting alpha to 0, runs Ridge.

plot(Ridgecv) #lowest MSE is when lambda = its lowest value. also variance is not that big at that value

#train model
fit.Ridge <- glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 0, family = 'gaussian', standardize = F,
                    lambda = Ridgecv$lambda.min)

#variable importance
ridgemat <- as.matrix(coef(fit.Ridge))
imp_Ridge <- names(ridgemat[abs(ridgemat) > 0.05,])[-1]

#evaluate model
pred_ridge <- predict(fit.Ridge, newx = test[,2:ncol(test)])
metrics_ridge <- metrics(test[,1], pred_ridge)

@@

###Elastic Net
```{r}
#to cross-validate alpha
alpha <- seq(0.1, 0.9, by = 0.1)

#cross-validation of alpha
set.seed(6164900)
glmnetcvs <- list()
for (i in alpha) {
    glmnetcvs[[which(alpha == i)]] <- cv.glmnet(x = train[, 2:ncol(train)], y = train[,1], family = 'gaussian',
                                              standardize = F, nfolds = K, alpha = i)
}
names(glmnetcvs) <- sprintf("alpha%g", alpha)

#obtain lowest lambda and check variance of model at that lambda by looking at the plot
minlambda <- sapply(glmnetcvs, function(x){
  plot(x)
  x$lambda.min
})
min.lambda <- minlambda[minlambda == min(minlambda)] #alpha = 0.9, lambda = 0.0029

#train model
fit.EN <- glmnet(x = train[, 2:ncol(train)], y = train[,1], alpha = 0.9, family = 'gaussian', standardize = F,
                    lambda = min.lambda)

#variable importance
ENmat <- as.matrix(coef(fit.EN))
imp_EN <- names(ENmat[abs(ENmat) > 0.05,])[-1]

#evaluate model
pred_EN <- predict(fit.EN, newx = test[,2:ncol(test)])
metrics_EN <- metrics(test[,1], pred_EN)

###LPCR
preprocess for LPCR
```{r}
train <- cbind(train, ind = sample(rep(1:K, length.out = nrow(train)))) #indicator variable
traindf <- as.data.frame(train)
testdf <- as.data.frame(test)

Linear Principle Component regression algorithm.
```{r}
npc <- 1:4 #hyperparameters to be cross-validated

metrics_lpcr <- list() #save results of each condition in a list
for(j in npc){ #for every condition
  metric_cv <- matrix(NA, K, 3) #save results for a fold in matrix
  colnames(metric_cv) <- c("RMSE", "MAD", "R2") #give comprehensive names
  for(i in 1:K){ #for everly fold
    traincv <- traindf[traindf$ind != i,] #set training set
    validation <- traindf[traindf$ind == i,] #set testing set

   metric_cv[i,] <- LPCR(traincv, validation, npc = j, relevance = F) #run LPCR to obtain only metrics and save
  }
  metrics_lpcr[[j-(npc[1]-1)]] <- metric_cv #save in storage
}

names(metrics_lpcr) <- sprintf("NPC%d", npc) #give comprehensive names

LPCR obtain best condition
```{r}
eval.pcr <- evaluation(metrics_lpcr) #evaluate the list of conditions
eval.pcr$evaluation #list of metrics per condition
eval.pcr$RMSE #not much difference between 3 and 4
eval.pcr$lowest_RMSE #using 4 PC's gives the lowest RMSE, however it is not much better than when 3 PC's are used, so 3 is preferred. 
```

LPCR test dataset
```{r}
fit_lpcr <- LPCR(traindf, testdf, npc = 3, relevance = T) #set relevance to true returns variable importance
metrics_lpcr <- fit_lpcr$metrics
imp_lpcr <- fit_lpcr$imp.pcr
```

###obtain final results
```{r}
#importance_lst <- ls(pattern = "imp_")
#metrics_lst <- ls(pattern = "metrics_")
#remove all variables except the ones we still need
#rm(list=setdiff(ls(), c(importance_lst, metrics_lst)))

#obtain metrics
result_metrics <- do.call(rbind, mget(ls(pattern = "metrics_")))

#obtain variable importance
result_feature_importance <- table(do.call(c, mget(ls(pattern = "imp_"))))

results <- list(performance = result_metrics, feature_imp = result_feature_importance)

saveRDS(results, "../Results/Results.RData")

###delete environment and console if desired
```{r}
rm(list=ls())
cat("\014")
```
